SETI in the Spatio-Temporal Survey Domain \\ {\color{red} {\bf DRAFT} }

### 2019-05-29 paper/ms.tex
finished edits

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%
\title{SETI in the Spatial-Temporal Survey Domain \\
{\color{red} {\bf DRAFT} }
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Traditional searches for extraterrestrial intelligence (SETI) or ``technosignatures'' focus on dedicated observations of single stars or regions in the sky to detect excess or transient emission from intelligent sources. The newest generation of synoptic time domain surveys enable an entirely new approach: spatial--temporal SETI, where technosignatures may be discovered from spatially resolved sources or multiple stars over time. 
Current optical time domain surveys such as ZTF and the Evryscope can probe 10--100 times more of the ``Cosmic Haystack'' parameter space volume than many radio SETI investigations. Small-aperture, high cadence surveys like Evryscope can be comparable in their Cosmic Haystack volume completeness to deeper surveys including LSST. Investigations with these surveys can also be conducted at a fraction of the cost of dedicated SETI surveys, since they require analysis of data already being gathered. 
However, SETI methodology has not widely utilized such surveys, and the literature is in need of new search algorithms that can incorporate signals from both the spatial and temporal domains.
Here I describe the broad potential for modern wide-field time domain optical surveys to revolutionize our search for technosignatures, and illustrate some example SETI approaches using transiting exoplanets to form a distributed beacon.
\end{abstract}

@@

Despite more than half a century of activity, the Search for Extra-Terrestrial Intelligence (SETI) is a field of study very much in its infancy. 
\citet{wright2018c} note the volume of parameter space for SETI (or the search for a needle amongst the ``cosmic haystack'') at radio wavelengths has been barely explored, and that many obvious signals may indeed be awaiting our discovery. 
The lack of SETI activity by the astronomical community can be partially understood as due to funding limitations in recent decades, as well as social pressures against studying SETI experienced by professional astronomers and organizations \citep{wright2018b}.

Traditional SETI work also requires time-consuming observations, often through dedicated monitoring of nearby stars by radio telescopes. Such observing campaigns are expensive and difficult to obtain given the competitive nature of telescope allocation. Recent progress for systematic ``technosignature'' searches has been made by Breakthrough Listen \citep{worden2017,isaacson2017}, primarily at radio wavelengths \citep[e.g.][]{price2018}.

The SETI conundrum has always been not only ``{\it What should we look for?}'', but  also {\it ``When, where, and how to look?}''. 
While \citet{wright2018c} suggest the completeness of our search of the ``cosmic haystack'' at radio wavelengths is akin to the ratio of a swimming pool compared to the Earth's ocean's ($\sim$10$^{-20}$), SETI at optical wavelengths is surely many orders of magnitude less complete. However, wide-field time domain surveys enable us to greatly expand our search in two key dimensions of parameter space: 1) collectively observing nearly the entire night sky, and 2) providing time resolved monitoring over many years.

\citet{djorgovski2000} highlighted the opportunity for a new age of generalized SETI using a virtual observatory made of digital sky surveys.
New missions like the currently running Zwicky Transient Facility \citep[ZTF][]{bellm2014} and the upcoming Large Synoptic Survey Telescope \citep[LSST][]{lsst} will be transformative in enabling both reproducible and cost-efficient SETI. By developing search strategies that utilize public survey databases and real-time ``alert streams'', both of which are being developed to facilitate a wide range of science goals from these facilities, optical survey SETI can be conducted automatically. As new search algorithms are developed, these databases can be quickly re-analyzed, providing a much-needed level of reproducibility and transparency that will help reduce the ``giggle factor'' stigma for SETI \citep{wright2018b}.

In this paper I explore the potential for wide-field time domain surveys in conducting a new generation of technosignature searches, starting with an overview of the motivation for doing SETI with optical surveys in \S\ref{sec:method}.
In \S\ref{sec:strategy} I describe the general principles behind technosignature searches in time domain surveys, including spatial over-densities of light curve signals and coordinated signals between multiple sources, for example.
In \S\ref{sec:haystack} I introduce several relevant surveys, and use the ``Cosmic Haystack'' search volume metric developed by \citet{wright2018c} to quantitatively compare these surveys to current radio SETI approaches. 
The data-driven spatial-temporal domain explored by these surveys necessitates new SETI approaches be developed, and in \S\ref{sec:transit} I outline an example signal where transiting planets around many stars could act as a spatially distributed beacon.
Finally in \S\ref{sec:discussion} I discuss ideas for other search approaches, and the advantages in cost and reproducibility of conducting SETI with current and upcoming optical time domain surveys.

@@

% The first SETI challenges: WHEN, WHERE, WHAT to search
Wide-field imaging surveys, and their associated infrastructure, are an ideal data source for carrying out technosignature searches that augment efforts in other wavelengths or with different observing strategies. By covering large portions of the night sky with repeated imaging, and with surveys that span years to decades in duration, we can substantially reduce the challenge of deciding {\it when} and {\it where} to look. Increasingly, multiple surveys obtain imaging that overlaps in both time and sky coverage, but with complimentary properties (e.g. imaging depth, cadence, wavelength coverage), which allows for rapid vetting and characterization of rare or low signal-to-noise events.. This also results in survey data providing some ``ground truth'' about the nature of most known objects in our Galaxy (e.g. their positions, temperatures, etc), and allows us to eliminate uncertainty on {\it what} objects we're studying for SETI.

% next: HOW to search (can compare approaches, and make them reproducible)
Optical surveys are also an excellent platform for exploring {\it how} to conduct technosignature searches. The large archives of data produced from these surveys become valuable legacies that enable science long after data acquisition has concluded. Searches for new phenomena in time-domain astronomy can be carried out both in the future with new observations and in the past by re-analyzing survey archives. 
Using survey archives for SETI as opposed to targeted observations allows us to compare the results of competing algorithms or strategies, and enables reproducible science. Search algorithms are also portable, and ideally can be used to carry out SETI on any appropriately structured large survey.
As future survey data is gathered, new classes of technosignature signals may be detectable that rely on decades worth of data. 
Since any candidate technosignature detection will (or should) be the subject of great scrutiny, making the data and algorithms open and reproducible will be critical for comparison and independent validation.

% DO IT FOR "FREE"
Large surveys, particularly ground-based efforts like LSST, are a highly prioritized component of the current and future development landscape in astronomy \citep{astro2010}. As such, great efforts are being made to develop survey tools and technologies to enable science from these missions, such as new database technologies \citep{juric2012}. These developments are especially needed to utilize the new forms of data being made available, including real-time data analysis \citep[e.g.][]{schwamb2019} and broadcasting real-time alerts for events \citep{patterson2019}.
The infrastructure needed to conduct technosignature searches with these large optical surveys is therefore already being developed. As with many areas of astrophysics in the ``survey era'', the biggest restriction to SETI with large surveys is therefore the availability of supported researchers mine the data.

@@

Here I describe a few general classes of technosignature signals that may be well suited to developing new detection algorithms for spatial-temporal surveys:

\begin{enumerate}
\item {\bf Unusual variability profiles or statistical distributions of fluxes }-- either on short timescales such as Boyajian's Star \citep{boyajian2015}, or long timescales such as disappearing stars \citep{villarroel2016} or occasionally missing transits \citep{kipping2016}
%
\item {\bf Spatial correlations of events or phenomena }-- e.g. coordination of transiting systems, or rebroadcasting events such as novae along the ``SETI ellipse'' \citep{makovetskii1977,lemarchand1994,tarter2001,shostak2004}
%
\item {\bf Spatial over-density or distribution} -- e.g. an over-density of a given phenomena, such as within the Earth Transit Zone \citep{heller2016}, or in spatial clusters
%
\item {\bf Unnatural patterns} -- particularly patterns of otherwise normal astrophysical variability, such as broadcasting a prime or fibonacci sequence of transits \citep{arnold2005,wright2016}.
%
\end{enumerate}

@@

\citet{wright2018c} have provided a framework to quantify the volume of observable parameter space covered by various SETI projects -- dubbed the ``Cosmic Haystack''. The properties for several major radio SETI projects yield a Haystack volume coverage of $10^{-20}$ to $10^{-18}$, or as \citet{wright2018c} illustrate, akin to a drinking glass of water compared to the volume of Earth's oceans. While this tool was developed with radio surveys in mind, it is broadly applicable to surveys at other wavelengths that are not explicitly designed for SETI \citep[e.g. see][]{forgan2019}.

While the potential for optical time domain surveys to advance technosignature searches has been qualitatively laid out in \S\ref{sec:method}, it is important to quantitatively compare them to traditional searches in the radio. 
For typical optical surveys using broadband photometry \citep{djorgovski2013}, the key parameters that determine the Cosmic Haystack volume coverage are the sky coverage ($\Omega$ in sq deg), total integration time per target ($T$ in sec), and survey sensitivity or per-epoch depth ($S$ in Jy).
Since these terms all act as scalers in the Haystack volume integration, surveys with different properties (e.g. few deep exposures versus many shallow exposures) can have comparably large Haystack coverage, filling out very different portions of the search volume. For maximum impact, SETI should therefore be conducted on many types of surveys.
%we need to be looking at big optical surveys, since they provide such amazing sky/time coverage

As a demonstration, I have computed the Cosmic Haystack volume coverage for five representative current and future optical time domain surveys, as shown in Figure \ref{fig:hay}. This includes the Catalina Real Time Transient Survey \citep[CRTS;][]{catalina}, Zwicky Transient Facility \citep[ZTF;][]{bellm2014}, the Transiting Exoplanet Survey Satellite \citep[\TESS;][]{tess}, the Large Synoptic Survey Telescope \citep[LSST;][]{lsst}, and the Evryscope \citep{law2015}. These include space-based and traditional ground-based surveys, and vary greatly in telescope collecting area, total numbers of visits, and survey footprint design. 
I have assumed sensitivity ($S$) is the published per-visit photometric depth for each survey. The total integration time per target ($T$) was computed as the average number of visits per target in the survey multiplied by the per-visit exposure time. The same power of transmitter is assumed as in \citet{wright2018c}. 
The code to generate the figures in this manuscript is online.\footnote{\url{https://github.com/TheAstroFactory/survey_seti}}

@@

Typical values for the Haystack fraction for radio SETI programs as presented in \citet{wright2018c}  are in the range of $\log V_{Haystack} = $ 20 --17. However, these programs are mostly targeted observations of nearby stars. The largest program featured in \citet{wright2018c}, by \citet{tingay2018}, had $\Omega=400$ sq deg and $T=10800$ seconds, and yielded $\log V_{Haystack} =  -17.36$.

By comparison, the wide field optical surveys I have featured all out perform the radio programs explored by \citet{wright2018c} in their Haystack fraction.
Due to its very wide field coverage, and near continuous monitoring each night, Evryscope (abbreviated as ``Evry'' in Figure \ref{fig:hay}) ranks as the best optical survey for technosignature searches of those surveyed here. The remarkable similarity in Haystack fraction between Evryscope, which uses 7-cm diameter telescopes, and the 8.4-m LSST survey reinforces the need to sample the Cosmic Haystack volume with as many different survey ``shapes'' as possible.

\citet{wright2018c} likened their Haystack fraction to searching for fish in a drinking glass of water compared to the Earth's oceans,. Our results for SETI with optical surveys are closer to 1-2 Olympic sized swimming pools (each 2500 m$^3$, according FINA standards) compared to the Earth's oceans\footnote{Calculation done via Wolfram$|$Alpha}. While optical surveys still only cover a tiny fraction of the total volume of parameter space, it is a substantial improvement.

@@

To illustrate the potential for possible technosignature searches in the spatial-temporal survey domain, in this section I demonstrate two examples of signals that rely on samples of transiting exoplanets. Transits are a promising target for SETI since they are relatively efficient to find over wide areas with surveys like \Kepler \citep{borucki2010} or \TESS, and they represent a natural ``Schelling point'' to attract attention with \citep{wright2017}. Creating such a beacon requires either generating a transiting system with unusual properties or patterns \citep[e.g.][]{arnold2005,kipping2016,forgan2017}, or as outlined in \S\ref{sec:strategy} an unusual {\it distribution} of transiting systems. The two examples outlined here explore the latter scenario, though I note the former is also well suited to automated approaches that would make use of future surveys and should be developed further. While I do not advocate these specific metrics are ideal for SETI, they demonstrate the type of signals we should design software systems to find in modern surveys.

@@

The easiest conceptual method to search for a technosignature beacon in the spatial distribution of exoplanets would be looking for over-densities, i.e. regions with an unusually large number of transiting systems. However this requires a large and very uniform sampling of exoplanets per square degree to determine if the density is unusual. Such spatially uniform samples are difficult to achieve since e.g. bright stars can saturate large regions.

@@

In Figure \ref{fig:2pt} I show the two-point correlation function up to spatial scales of 5 degrees for both the sample of 2,324 transiting exoplanets discovered by the \Kepler primary mission,\footnote{Catalog from \url{http://exoplanetarchive.ipac.caltech.edu}} and for 201,312 stars in the \Kepler field. Since the sample of exoplanets was relatively small, uncertainties shown in Figure \ref{fig:2pt} were estimated by bootstrapping the correlation function 500 times. Here the correlation was done using only the (ra,dec) coordinates for each star or exoplanet system, but can also be done for additional dimensions (e.g. including distance, orbital properties, etc).

@@

Rather than searching for an exoplanet beacon consisting of outliers in spatial density or distribution (e.g. \S\ref{sec:2pt}), we can consider groups of systems that are unusually similar in some way. In the simplest case this could take the form of many closely clustered stars with transiting exoplanets, where each planet had the same radius, transit depth, transit ephemeris (or transit midpoint), or orbital period, for example. The assumption here is that a civilization would have engineered this unlikely set of transiting systems, like a lighthouse, in order to be noticed by early astronomers (us) hoping to characterize the exoplanet demographics of our Galaxy. This SETI approach relies on both the spatial and temporal information contained in exoplanet-hunting surveys like \Kepler or \TESS.
Here I demonstrate a simple algorithm for searching for such clusters of exoplanets with unusually similar orbital periods.

%two questions arise: 
%- how many systems to use as a cluster, and how spatially tight should they be?
%- which periods and how self-similar do they need to be?

@@

For each $k=3$ star cluster I compute two metrics that should be considered when searching for  spatially distributed technosignature beacons, shown in Figure \ref{fig:knn}. First I simply calculate the mean cartesian distance of the three exoplanet systems about the ``center'' or average spatial position. This enables us to potentially identify clusters of transiting systems that are unusually close together in projected separation. Second I compute the similarity of the orbital periods for the three exoplanets in each cluster. This similarity metric here is defined as the range (i.e. max period -- min period) divided by the mean orbital period of the three stars.
This period metric is easy to interpret, allowing us to identify clusters of transiting exoplanets whose orbital periods deviate by less than a given percentage.

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{../figures/Kplanets.pdf}
\caption{Fractional range of orbital periods versus the mean angular separation for KNN clusters in (ra, dec) space of transiting exoplanet systems in \Kepler. This sample only includes systems with only a single known transiting exoplanet. These KNN clusters have typical spatial separations of $\sim$0.1 deg. For $K$=3, three unique clusters have a spread in their orbital periods below 6\% (red points). However this is not a significant outlier in the orbital period similarity, and these clusters are not unusually close together in their spatial separation.}
\label{fig:knn}
\end{figure}

@@

In this paper I have presented the great potential for using algorithms to search for SETI or ``technosignature'' signals in modern spatial--temporal surveys, such as TESS, ZTF, or LSST. Though these surveys are primarily in optical wavelengths as opposed to traditional SETI at radio frequencies, there are many promising algorithmic approaches that may be considered, and these searches can ``piggyback'' on datasets and computational resources being developed for other science goals.
This SETI approach also connects with the longterm rise of data science and statistical methods in astronomy to characterize, and find outliers in, large datasets. Interestingly, the search for technosignatures is possibly one of only a few areas where ``big data'' survey astronomy intersects with the growing astrobiology discipline. As \citet{berea2019} note, expertise in many complimentary domains of data analysis (e.g. cybersecurity) may also be useful for advancing survey SETI.

@@

As described in the introduction, these sky surveys enable a wide range of algorithmic approaches to use for SETI. Real-time alert streams from surveys like ZTF and soon LSST \citep[e.g.][]{schwamb2019} make rapid identification and follow-up of unusual variability straightforward to implement. This could include monitoring for over-densities in the spatial distribution of alerts, especially within the Earth Transit Zone \citep{heller2016}, for slowly appearing or disappearing stars on decades timescales \citep{villarroel2016}, or for coordinated variability with e.g. novae along the ``SETI ellipsoid'' over wide areas of the sky \citep{lemarchand1994}. In figure work we will explore a framework for searching the ZTF alert stream for unusual or ``impossible'' sequences of events.

@@

@@

J.R.A.D. thanks Nicholas M. Law and Jason T. Wright for their helpful ongoing conversations in developing this idea, and Andrew P. V. Siemion and the Berkeley SETI Research Center team for hosting me for an enlightening and motivating visit, and the Research Corporation for Scientific Advancement for hosting the 2018 TDA Scialog meeting.

>>>>

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%
\title{SETI in the Spatio-Temporal Survey Domain \\
{\color{red} {\bf DRAFT} }
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Traditional searches for extraterrestrial intelligence (SETI) or ``technosignatures'' focus on dedicated observations of single stars or regions in the sky to detect excess or transient emission from intelligent sources. The newest generation of synoptic time domain surveys enable an entirely new approach: spatio--temporal SETI, where technosignatures may be discovered from spatially resolved sources or multiple stars over time. 
Current optical time domain surveys such as ZTF and the Evryscope can probe 10--100 times more of the ``Cosmic Haystack'' parameter space volume than many radio SETI investigations. Small-aperture, high cadence surveys like Evryscope can be comparable in their Haystack volume completeness to deeper surveys including LSST. Investigations with these surveys can also be conducted at a fraction of the cost of dedicated SETI surveys, since they make use of data already being gathered. 
However, SETI methodology has not widely utilized such surveys, and the field is in need of new search algorithms that can account for signals in both the spatial and temporal domains.
Here I describe the broad potential for modern wide-field time domain optical surveys to revolutionize our search for technosignatures, and illustrate some example SETI approaches using transiting exoplanets to form a distributed beacon.
\end{abstract}

@@

Despite more than half a century of activity, the Search for Extra-Terrestrial Intelligence (SETI) -- also described as the search for ``technosignatures'' -- is a field of study very much in its infancy. 
The lack of SETI research by the astronomical community is partially due to funding limitations in recent decades, as well as social pressures against studying SETI experienced by professional astronomers and organizations \citep{wright2018b}. 
Indeed, the volume of parameter space for SETI (or the search for a needle amongst the ``Cosmic Haystack'') at radio wavelengths has been barely explored \citet{wright2018c}, and thus many obvious signals may be awaiting our discovery. 
However, the new era of surveys and data-driven discovery in astronomy holds great promise for revolutionizing SETI.

Traditional SETI work requires time-consuming observations, often through dedicated monitoring of nearby stars by radio telescopes. Such observing campaigns are expensive and difficult to obtain given the competitive nature of telescope allocation. Recent progress for systematic ``technosignature'' searches has been made by Breakthrough Listen \citep{worden2017,isaacson2017}, primarily at radio wavelengths \citep[e.g.][]{price2018}.

The SETI conundrum has always been not only ``{\it What should we look for?}'', but  also {\it ``When, where, and how}''. 
While \citet{wright2018c} suggest the completeness of our search of the ``Cosmic Haystack'' at radio wavelengths is akin to the ratio of a small swimming pool compared to the Earth's oceans ($\sim$10$^{-20}$), SETI at optical wavelengths is surely many orders of magnitude less complete.  
Current optical-SETI is typically focused on searches for nanosecond pulses with specialized wide-field high-speed imaging \citep{howard2004}, or unusual spectroscopic features \citep{isaacson2019}.
Wide-field time domain photometric surveys enable us to greatly expand our search in two key dimensions of parameter space: 1) collectively observing nearly the entire night sky, and 2) providing precision time-resolved monitoring over many years.

\citet{djorgovski2000} highlighted the opportunity for a new age of generalized SETI using a virtual observatory made of digital sky surveys.
New missions like the currently running Zwicky Transient Facility \citep[ZTF;][]{bellm2014} and the upcoming Large Synoptic Survey Telescope \citep[LSST;][]{lsst} will be transformative in enabling both reproducible and cost-efficient SETI. By developing search strategies that utilize public survey databases and real-time ``alert streams'', both of which are being developed to facilitate a wide range of science goals from these facilities, optical survey SETI can be conducted automatically. As new search algorithms are developed, these databases can be re-analyzed, providing a much-needed level of reproducibility and transparency that will help reduce the ``giggle factor'' stigma for SETI \citep{wright2018b}.

In this paper I explore the potential for wide-field time domain surveys in conducting a new generation of technosignature searches, starting with an overview of the motivation for doing SETI with optical surveys in \S\ref{sec:method}.
In \S\ref{sec:strategy} I describe the general principles behind technosignature searches in time domain surveys, including spatial over-densities of light curve signals and coordinated signals between multiple sources, for example.
In \S\ref{sec:haystack} I introduce several relevant surveys, and use the ``Cosmic Haystack'' search volume metric developed by \citet{wright2018c} to quantitatively compare these surveys to current radio SETI approaches. 
The data-driven spatio-temporal domain explored by these surveys necessitates new SETI approaches be developed, and in \S\ref{sec:transit} I outline an example signal where transiting planets around many stars could act as a spatially distributed beacon.
Finally in \S\ref{sec:discussion} I discuss ideas for other search approaches, and the advantages in cost and reproducibility of conducting SETI with current and upcoming optical time domain surveys.

@@

% The first SETI challenges: WHEN, WHERE, WHAT to search
Wide-field imaging surveys, and their associated infrastructure, are an ideal data source for carrying out technosignature searches that augment efforts in other wavelengths or with different observing strategies. By covering large portions of the night sky with repeated imaging, and with surveys that span years to decades in duration, we can substantially reduce the challenge of deciding {\it when} and {\it where} to look. Increasingly, multiple surveys obtain imaging that overlaps in both time and sky coverage, but with complementary properties (e.g. imaging depth, cadence, wavelength coverage), which allows for rapid vetting and characterization of rare or low signal-to-noise events. Surveys also produce important baseline information about the known objects in our Galaxy (e.g. their positions, temperatures, etc), which allows us to eliminate uncertainty on {\it what} objects being targeted for SETI.

% next: HOW to search (can compare approaches, and make them reproducible)
Optical surveys are also an excellent platform for exploring {\it how} to conduct technosignature searches. The large archives of data produced from these surveys become valuable legacies that enable science long after data acquisition has concluded. Searches for new phenomena in time-domain astronomy can be carried out both in the future with new observations and in the past by re-analyzing survey archives. 
Using survey archives for SETI as opposed to targeted observations of individual stars allows us to compare the results of competing algorithms or strategies, and enables reproducible science. Search algorithms are also portable, and ideally can be used to carry out SETI on any new survey or large ensemble of observations.
As future survey data is gathered, new classes of technosignature signals may be detectable that rely on decades worth of data. 
Since any technosignature detection candidate should be the subject of great scrutiny, making the data and algorithms open and reproducible will be critical for comparison and independent validation.

% DO IT FOR "FREE"
Large surveys, particularly ground-based efforts like LSST, are a highly prioritized component of the current and future development landscape in astronomy \citep{astro2010}. As such, great efforts are being made to develop survey tools and technologies to enable science from these missions, such as new database technologies \citep{juric2012,zecevic2019}, real-time event broadcasting \citep{patterson2019}, and real-time data analysis frameworks \citep[e.g.][]{schwamb2019}.
The infrastructure needed to conduct technosignature searches with these large optical surveys is therefore already being developed. As with many areas of astrophysics in the ``survey era'', the biggest bottleneck for SETI with these surveys is the support (both financial and social) for researchers to develop algorithms and mine the available data.

@@

Here I describe a few general classes of technosignature signals that may be well suited to developing new detection algorithms for spatio-temporal surveys:

\begin{enumerate}
\item {\bf Unusual variability profiles or statistical distributions of fluxes }-- either on short timescales such as Boyajian's Star \citep{boyajian2015}, or long timescales such as disappearing stars \citep{villarroel2016} or occasionally missing transits \citep{kipping2016}
%
\item {\bf Spatial correlations of events or phenomena }-- e.g. coordination of transiting systems, or rebroadcasting events such as novae along the ``SETI ellipse''  that defines the ellipsoid for receiving synchronized signals from multiple transmitters 
\citep{makovetskii1977,lemarchand1994,tarter2001,shostak2004}
%
\item {\bf Spatial over-density or distribution} -- e.g. an over-density of a given phenomena, such as within the ``Earth Transit Zone'' band of stars that would see Earth as a transiting exoplanet \citep{heller2016}, or in spatial clusters
%
\item {\bf Unnatural patterns} -- particularly patterns of otherwise normal astrophysical variability (e.g. flares, pulsations, transits, etc), such as broadcasting a prime or fibonacci sequence using transits \citep{arnold2005,wright2016}.
% 
\end{enumerate}
% omit piloted spacecraft, e.g. like looking for O'muamua objects

@@

While the potential for optical time domain surveys to advance technosignature searches has been qualitatively laid out in \S\ref{sec:method}, it is important to quantitatively compare them to traditional searches in the radio. 
\citet{wright2018c} have provided a framework to quantify the volume of observable parameter space covered by various SETI projects -- dubbed the ``Cosmic Haystack''. Though this metric was developed with radio surveys in mind, it is broadly applicable to surveys at other wavelengths that are not explicitly designed for SETI \citep[e.g. see][]{forgan2019}.

For typical optical surveys using broadband photometry \citep{djorgovski2013}, the key parameters that determine the Haystack volume coverage are the sky coverage ($\Omega$ in sq deg), total integration time per target ($T$ in sec), and survey sensitivity or per-epoch depth ($S$ in Jy).
Since these terms all act as scalers in the Haystack volume integration, surveys with different properties (e.g. few deep exposures versus many shallow exposures) can have comparably large Haystack coverage, filling out very different portions of the search volume. For maximum impact, SETI should therefore be conducted on many types of surveys.

As a demonstration, I have computed the Haystack volume coverage for five representative current and future optical time domain surveys, as shown in Figure \ref{fig:hay}. This includes the Catalina Real Time Transient Survey \citep[CRTS;][]{catalina}, Zwicky Transient Facility \citep[ZTF;][]{bellm2014}, the Transiting Exoplanet Survey Satellite \citep[\TESS;][]{tess}, the Large Synoptic Survey Telescope \citep[LSST;][]{lsst}, and the Evryscope \citep{law2015}. These include space-based and ground-based surveys, and vary greatly in telescope collecting area, total numbers of visits, and survey footprint design. 
I have assumed sensitivity ($S$) is the published per-visit photometric depth for each survey. The total integration time per target ($T$) was computed as the average number of visits per target in the survey multiplied by the per-visit exposure time. The same power of transmitter is assumed as in \citet{wright2018c}. 
The code to generate the figures in this manuscript is online.\footnote{\url{https://github.com/TheAstroFactory/survey_seti}}

@@

Typical values for the Haystack fraction for radio SETI programs as presented in \citet{wright2018c}  are in the range of $\log V_{Haystack} =  -20$ to $-17$. However, these programs are mostly targeted observations of nearby stars. The largest program featured in \citet{wright2018c}, by \citet{tingay2018}, had $\Omega=400$ sq deg and $T=10800$ seconds, and yielded $\log V_{Haystack} =  -17.36$.

The wide field optical surveys I have examined here all out perform the radio programs explored by \citet{wright2018c} in their Haystack fraction. Due to its very wide field coverage, and near continuous monitoring each night, Evryscope (abbreviated as ``Evry'' in Figure \ref{fig:hay}) ranks as the best optical survey for technosignature searches of those surveyed here. The remarkable similarity in Haystack fraction between Evryscope, which uses 7-cm diameter telescopes, and the 8.4-m LSST survey reinforces the need to sample the Haystack volume with as many different survey ``shapes'' as possible. Note that since the area coverage, photometric depth, and cadence of these two surveys are so different, they will each be sensitive to very different specific SETI detection properties, such as signal repetition timescales or amplitudes.

\citet{wright2018c} likened their Haystack fraction to searching for fish in a hot tub or small swimming pool compared to the Earth's oceans. Our results for SETI with optical surveys are closer to 1-2 Olympic sized swimming pools (each 2500 m$^3$, according FINA standards) compared to the Earth's oceans\footnote{Calculation done via Wolfram$|$Alpha}. While optical surveys still only cover a tiny fraction of the total volume of parameter space, it is a substantial improvement.

@@

To illustrate the potential for possible technosignature searches in the spatio-temporal survey domain, in this section I demonstrate two examples of signals that rely on samples of transiting exoplanets. Transits are a promising target for SETI since they are relatively efficient to find over wide areas with surveys like \Kepler \citep{borucki2010} or \TESS, and they represent a natural object to attract attention with, known as a ``Schelling point'' \citep{wright2017}. Creating such a beacon requires either generating a transiting system with unusual properties or patterns \citep[e.g.][]{arnold2005,kipping2016,forgan2017}, or as outlined in \S\ref{sec:strategy} an unusual {\it distribution} of transiting systems. The two examples outlined here explore the latter scenario, though I note the former is also well suited to automated approaches that would make use of future surveys and should be developed further. While I do not advocate these {\it specific} metrics are ideal for SETI, they demonstrate the type of signals we should design software systems to find in modern surveys.

@@

The easiest conceptual method to search for a technosignature beacon in the spatial distribution of exoplanets would be looking for over-densities, i.e. regions with an unusually large number of transiting systems relative to the local density of stars. However this requires a large and very uniform sampling of exoplanets per square degree to determine if the density is unusual. Such spatially uniform samples are difficult to achieve since e.g. bright stars can saturate large regions.

@@

In Figure \ref{fig:2pt} I show the two-point correlation function up to spatial scales of 5 degrees for both the sample of 2,324 transiting exoplanets discovered by the \Kepler primary mission,\footnote{Catalog from \url{http://exoplanetarchive.ipac.caltech.edu}} and for 201,312 stars in the \Kepler field. Since the sample of exoplanets was relatively small, uncertainties shown in Figure \ref{fig:2pt} were estimated by bootstrapping the correlation function 500 times. Here the correlation was done using only the (RA, Dec) coordinates for each star or exoplanet system, but can also be done for additional dimensions (e.g. including distance, orbital properties, etc).

@@

Rather than searching for an exoplanet beacon consisting of outliers in spatial density or distribution (e.g. \S\ref{sec:2pt}), we can consider groups of systems that are unusually similar in some way. In the simplest case this could take the form of many closely clustered stars with transiting exoplanets, where each planet had the same radius, transit depth, transit ephemeris (or transit midpoint), or orbital period, for example. The assumption here is that a civilization would have engineered this unlikely set of transiting systems, like a lighthouse, in order to be noticed by early astronomers who are characterizing the exoplanet demographics of our Galaxy (i.e. us). This SETI approach relies on both the spatial and temporal information contained in exoplanet-hunting surveys like \Kepler or \TESS.
Here I demonstrate a simple algorithm for searching for such clusters of exoplanets with unusually similar orbital periods.

@@

For each $k=3$ star cluster I compute two metrics that should be considered when searching for  spatially distributed technosignature beacons, shown in Figure \ref{fig:knn}. First I simply calculate the mean cartesian distance of the three exoplanet systems about the ``center'' or average spatial position in 2D space (RA, Dec). This enables us to potentially identify clusters of transiting systems that are unusually close together in projected separation. Second I compute the similarity of the orbital periods for the three exoplanets in each cluster. This similarity metric here is defined as the range (i.e. max period -- min period) divided by the mean orbital period of the three stars.
This period metric is easy to interpret, allowing us to identify clusters of transiting exoplanets whose orbital periods deviate by less than a given percentage.

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{../figures/Kplanets.pdf}
\caption{Fractional range of orbital periods versus the mean angular separation for KNN clusters in (RA, Dec) space of transiting exoplanet systems in \Kepler. This sample only includes systems with only a single known transiting exoplanet. These KNN clusters have typical spatial separations of $\sim$0.1 deg. For $K$=3, three unique clusters have a spread in their orbital periods below 6\% (red points). However this is not a significant outlier in the orbital period similarity, and these clusters are not unusually close together in their spatial separation.}
\label{fig:knn}
\end{figure}

@@

In this paper I have presented the great potential for using algorithms to search for SETI or ``technosignature'' signals in modern spatio--temporal surveys, such as TESS, ZTF, or LSST. Though these surveys are primarily in optical wavelengths as opposed to traditional SETI at radio frequencies, there are many promising algorithmic approaches that may be considered. Further, these searches can ``piggyback'' on datasets and computational resources being developed for other science goals.
This SETI approach also connects with the rise of data science and statistical methods in astronomy to characterize, and find outliers in, large datasets. Interestingly, the search for technosignatures is possibly one of only a few areas where ``big data'' survey astronomy intersects with the growing astrobiology discipline. As \citet{berea2019} note, expertise in many complementary domains of data analysis (e.g. cybersecurity) may also be useful for advancing survey SETI.

@@

As described in the introduction, sky surveys enable a wide range of algorithmic approaches to use for SETI. Real-time alert streams from surveys like ZTF and soon LSST \citep[e.g.][]{schwamb2019} make rapid identification and follow-up of unusual variability straightforward to implement. Other approaches could include monitoring for over-densities in the spatial distribution of alerts, especially within the Earth Transit Zone \citep{heller2016}, searching for slowly appearing or disappearing stars on decades timescales \citep{villarroel2016}, or looking for coordinated variability with e.g. novae along the ``SETI ellipsoid'' over wide areas of the sky \citep{lemarchand1994}. Unnatural motions of objects could also be searched for in sky surveys. This might be akin to searching for interstellar asteroids \citep[e.g.][]{mamajek2017} whose trajectories suddenly change. In future work we will explore a framework for searching the ZTF alert stream for unusual or ``impossible'' sequences of events in light curves.

@@

\software{Python, IPython \citep{ipython}, NumPy \citep{numpy}, Matplotlib \citep{matplotlib}, SciPy \citep{scipy}, Pandas \citep{pandas}, Astropy \citep{astropy}}

@@

J.R.A.D. thanks Nicholas M. Law and Jason T. Wright for their ongoing conversations in developing this idea, Andrew P. V. Siemion and the Berkeley SETI Research Center team for hosting me for an enlightening and motivating visit, the Research Corporation for Scientific Advancement for hosting the 2018 TDA Scialog meeting, and Daniela Huppenkothen for comments that greatly improved this manuscript.
